\toolnameend: A Tool for Visualizing Patterns of Spreadsheet Function Combinations

### 2016-08-12 doc/paper.tex
Walkthrough changes

		\item [1] \textbf{Provide an interactive environment for free exploration.}
		There are a lot of data in some of these spreadsheet corpora, and the user
		should be able to pick for themselves which to explore while ignoring the
		rest. Furthermore, this exploration should not be designed around answering
		one specific question but rather helping to answer a variety of questions.

		\item [!1] \textbf{Do not make judgments about good and bad design.} Though
		users may certainly approach the tool with the intention of exploring
		particular formulae, a formula visualization tool will not prioritize any on
		metrics of good or bad practice.

		\item [2] \textbf{Emphasize the quantitative patterns in formula
			construction.} The tool should address the questions of how often the
		spreadsheet programmers used a certain function and where they used it. To do
		this, it must show metrics, such as frequency of use and depth of function
		nesting, of the dataset it conveys.

		\item [!2] \textbf{Do not overwhelm with numbers.} For popular functions
		especially, we anticipate that we'll have to show a lot of related data at
		once. Though everything should be accessible in accordance with Goal 1, we
		should design in a way that reduces noise.

		\item [3] \textbf{Promote a qualitative understanding of the patterns.} The
		tool should supplement its observations with concrete instances of relevant
		formulae from the corpus. Ideally, it should even direct users to the exact
		cell in the spreadsheets where the formula was used, contextualizing the
		functions.

		\item [!3] \textbf{Do not attempt to explain formulae.} Though the tool is
		meant to foster understanding by linking pattern to example, it will not try
		to be an explanatory tool that articulates what a complex formula
		accomplishes. Rather, we will provide the structure and context, and the user
		must infer this on their own.

@@

	Though we kept these guidelines in mind through the development of \toolname,
	we nevertheless reached a number of points in which these principles
	conflicted. For a discussion of these decisions, see Section
	\ref{ssec:decisions}.

	\subsection{Walkthrough of the Tool} \label{sec:walkthrough}Before discussing
	the minutiae of tool's development, we will give a basic example of how to use
	the tool. For this, we will work from the visualization built on the Enron
	corpus to answer this question: \textit{ What kinds of functions do people use
		to define the condition in an IF function?} \par

	First, a word on the IF function. In Excel, it serves as one of the logical
	functions: given a condition as the first argument that evaluates to either
	true or false, it takes on the value of either the second or third argument,
	respectively. Or, as the documentation writes out, ``IF(Something is True, then
	do something, otherwise do something
	else)"\footnote{https://support.office.com/en-us/article/IF-function-69aed7c9-4e8a-4755-a9bc-aa8bbff73be2?ui=en-US\&rs=en-US\&ad=US}. Furthermore, the third argument is optional; if left out, the function will return a default ``FALSE" value if necessary. Examples of this function in practice can be seen in Figure~\ref{fig:ifexample}.

@@

	When the user first approaches the tree for a given top-level function -- that
	is, a function nested within none other -- only a few nodes are visible, as
	shown in Figure \ref{fig:startpic}. \par

@@

	The visualization so far comprises two types of nodes: the circle
	\includegraphics{glossary-greenonly}, which represents a discrete function in
	the formula and is size according to its frequency relative to the number of
	times the root function appeared; and the numbered squares
	\includegraphics{glossary-blue}, which represent the positions of arguments
	within its parent function. From this, we can confirm that, of the times it was
	observed, IF can have at most three arguments passed into it, which corresponds
	with its specifications in the API. \par

>>>>

		\item [1] \textbf{Unbiased Exploration} First, provide an interactive
		environment for free exploration. There are a lot of data in some of these
		spreadsheet corpora, and the user should be able to pick for themselves which
		to explore while ignoring the rest. Second, avoid judgments about good and bad
		design. Though users may certainly approach the tool with the intention of
		exploring particular formulae, a formula visualization tool will not
		prioritize any on metrics of good or bad practice.

		\item [2] \textbf{Uncluttered Quantification} \toolname should emphasize the
		quantitative patterns in formula construction and address the questions of how
		often the spreadsheet programmers used a certain function and where they used
		it. To do this, it must show metrics, such as frequency of use and depth of
		function nesting, but it must not overwhelm with numbers. Though everything
		should be accessible in accordance with the first goal, we should reduce noise with our design.

		\item [3] \textbf{Informative Qualification} The tool should promote a qualitative understanding of the patterns by providing concrete instances of relevant
		formulae from the corpus. Ideally, it should even direct users to the exact
		cell in the spreadsheets where the formula was used, contextualizing the
		functions.

@@

	\subsection{Walkthrough of the Tool} \label{sec:walkthrough}Before discussing
	the tool's development, we will give a basic example of how to use
	the tool. For this, we will work from the visualization built on the Enron
	corpus to answer this question: \textit{ What kinds of functions do people use
		to define the condition in an IF function?} \par

	In Excel, the IF functions is one of the logical
	functions: given a condition as the first argument that evaluates to either
	true or false, it takes on the value of either the second or third argument,
	respectively. Or, as the documentation writes out, ``IF(Something is True, then
	do something, otherwise do something
	else)"\footnote{https://support.office.com/en-us/article/IF-function-69aed7c9-4e8a-4755-a9bc-aa8bbff73be2?ui=en-US\&rs=en-US\&ad=US}. Furthermore, the third argument is optional; if left out, the function will return a default ``FALSE" value if necessary. Examples of this function in practice can be seen in Figure~\ref{fig:ifexample}.

@@

	When the user first approaches the tree for a given top-level function -- that
	is, a function nested within none other -- only a few nodes are visible, as
	shown in Figure~\ref{fig:startpic}. \par

@@

	The visualization so far comprises two types of nodes: the circle
	\includegraphics{glossary-greenonly}, which represents a discrete function in
	the formula and is sized according to its frequency relative to the number of
	times the root function appeared; and the numbered squares
	\includegraphics{glossary-blue}, which represent the positions of arguments
	within its parent function. From this, we can confirm that, of the times it was
	observed, IF can have at most three arguments passed into it, which corresponds
	with its specifications in the API. \par
